---
title: "Test _R_^2^ calculation"
author: "TJ Mahr"
date: "July 26, 2017"
output: 
  rmarkdown::github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

This document is just my (tjmahr's) notes on the article... I have nothing to do
with its the paper besides being a studious reader.

*** 

I'm curious how the calculation of _R_^2^ compares to a function I wrote for my 
own analyses. 

This code is the function by the Stan Team.

```{r}
library(rstanarm)
library(ggplot2)
library(bayesplot)

# Compute Bayesian R-squared for linear and
# generalized linear models.
#
# @param fit A fitted model object returned by stan_glm.
# @return A vector of R-squared values with length equal to
#      the number of posterior draws.
#
bayes_R2 <- function(fit) {
  y <- get_y(fit)
  ypred <- posterior_linpred(fit, transform = TRUE)
  if (family(fit)$family == "binomial" && NCOL(y) == 2) {
    trials <- rowSums(y)
    y <- y[, 1]
    ypred <- ypred %*% diag(trials)
  }
  e <- -1 * sweep(ypred, 2, y)
  var_ypred <- apply(ypred, 1, var)
  var_e <- apply(e, 1, var)
  var_ypred / (var_ypred + var_e)
}
```

Here's the classical values.

```{r}
library(dplyr)

m_glm <- glm(weight ~ group, PlantGrowth, family = gaussian())
modelr::rsquare(m_glm, PlantGrowth)

m_lm <- lm(weight ~ group, PlantGrowth)
modelr::rsquare(m_lm, PlantGrowth)
```

Here's my calculation.

```{r}
m <- stan_glm(weight ~ group, PlantGrowth, family = gaussian(), prior = normal(0, 1))
m

naive_r2 <- tristan::calculate_model_r2(m)
tristan::tidy_etdi(naive_r2, c(.95, .9, .8))
tristan::tidy_median(naive_r2)

# Point value
tristan::calculate_r2(PlantGrowth$weight, predict(m))
```

Here is theirs.

```{r}
stanteam_r2 <- bayes_R2(m)
tristan::tidy_etdi(stanteam_r2, c(.95, .9, .8))
tristan::tidy_median(stanteam_r2)
```

This is kind of nutty.

```{r}
df <- list(naive = naive_r2,
  stanteam = stanteam_r2) %>% 
  tibble::enframe() %>% 
  tidyr::unnest(value)

library(ggplot2)

ggplot(df) + 
  aes(x = value) + 
  geom_histogram() + 
  theme_grey() + 
  facet_grid(name ~ .)
```

It looks like their median value is comparable to the modal _R_^2^ value in my 
implementation. That's nice... The median and interval stats in my
implementation are kind of worthless because the median _R_^2^ does not match
the _R_^2^ produced by the median model parameters..

To be clear, my naive implementation just does a really dumb calculation on each
posterior sample. Conceptually, each sample is its own lm model and each model
attains a certain _R_^2^ value so we have a distribution of _R_^2^ values.

```{r}
tristan::calculate_model_r2

tristan::calculate_r2

# To convince myself that it works
y_obs <- m_lm$model$weight
y_fits <- predict(m_lm)

tristan::calculate_r2(y_obs, y_fits)
modelr::rsquare(m_lm, PlantGrowth)

```

To do... study what their implementation does.
